---
title: "Check coordinate system"
author: "Stijn Van Hoey"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

## Introduction

When retrieving a number of data-points, but the kind of projection is not provided/known (for any reason), the `guess_projection` function plots the data for different projections on a map to make comparison possible. The main function is `guess_projection` which creates the map with different options. The `reproject_points` function can be useful as well, as it supports reprojection of coordinate columns in a `data.frame`.

## Aim

The aim is to quickly check the coordinate system by plotting the coordinates on a leaflet map with a list of given coordinate systems.

## Functionality

### Dependencies

In order to run the functionalities, some R packags need to be installed. The set of required libraries is handled by the `inborutils` package itself. 

Loading the functionality can be done by loading the `inborutils` package:

```{r}
library(inborutils)
```

### Projection function

The `guess_projection` function requires as minimal inputs:

1.  a `data.frame`
2.  the longitude column name
3.  the latitude column name. 

For example:

```{r fig.width=7, message=FALSE}
# my_dataframe has the columns "longitude" and "latitude"
guess_projection(coordinate_example, "longitude", "latitude")
```

Furthermore, the optional argument `belgium` is a boolean value (TRUE/FALSE) defining if the data is expectd to be in Belgium. Finally, the `epsg` codes of the corresponding coordinate systems to check can be customised. 

By default, the following CRS-systems are tested:

* `epsg:4326`
* `epsg:31370`
* `epsg:28992`
* `epsg:32631`
* `epsg:3812`
* `epsg:3035`

Which can be overridden by adapting the `projections` argument:

```{r fig.width=7, message=FALSE}
guess_projection(coordinate_example, "longitude", "latitude", belgium = TRUE, 
                 projections = c("epsg:2000", "epsg:2805", "epsg:4326"))
```

Check the [spatial reference](http://spatialreference.org) for the available codes.
