---
title: "INBOVEG queries as functions"
author: "Stijn Van Hoey"
date: "March 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(DBI)
opts_chunk$set(echo = TRUE)
```

# Introduction

In this tutorial we show how to use helper functions to query data directly from the INBOVEG SQL Server database. These functions wrap **SQL queries** and so avoid having to write the queries yourself or copy-paste queries generated with the access-frontend for INBOVEG.

We have provided functions to query information about surveys, metadata for relevés (= vegetation recording), the vegetation data and a query that helps to resolve taxonomic synonymy. 

# The database connection

In order to query data from INBOVEG, a database connection is required...

You have to make sure you are allowed to set up this connection. For more information, see [link to more general database connection tutorial](...)...

When the database name (the technical term is the [DSN](https://en.wikipedia.org/wiki/Data_source_name)) is known on your computer, you can make the connection as follows...

```{r connection, eval = FALSE}
con <- dbConnect(odbc::odbc(), dsn="Cydonia-prd") 
```

This connection will be the input argument for each of the INBOVEG query functions.


# Functionalities

### Survey overview

...

```{r eval = FALSE}
surveys <- inboveg_survey(con)
```


### Vegetation recording metadata

...

When interested in all surveys:

```{r eval = FALSE}
metadata <- inboveg_metadata_kop(con)
```

or focusing on a specific survey, for example `Sigma_LSVI_2012`:

```{r eval = FALSE}
metadata <- inboveg_metadata_kop(con, 
                                 survey_name = c("Sigma_LSVI_2012"))
```

### Coverage per vegetation layer

...



